---
title: "Week 1 - Basic Multivariate Summaries and Visualisations"
author: "Amanjit Gill"
date: "2022-10-26"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### **Task 1**

Basic data summaries.
```{r summaries1}
data(iris)

summary(iris)
```

Univariate summaries.
```{r summaries2}

# mean of each specified column
# last column, species, is discarded
colMeans(iris[,-5])

# apply sd(..) function to each column
# 2nd argument, 2, says to apply the function by column
apply(iris[,-5], 2, sd)

# apply quantile(.95) to each column
apply(iris[,-5], 2, quantile, .95)
```

Multivariate summaries.
```{r summaries3}
cov(iris[,-5])
cor(iris[,-5])

```

- Use 'pipe' operator %>% from the magrittr package to pass results from one function to the next.
- Use 'map' function from the purrr package.
```{r load_pipe}

library(magrittr)
library(purrr)
```

```{r use_pipe}

# this says:
# start with iris dataset
# split according to species column
# for each species, discard last column (species)
# compute covariances for the remaining columns, by species

iris %>%
    split(iris$Species) %>%
    map(~.[-5]) %>%
    map(cov)

```

#### **Task 2**

Visualisations.

- Use 'ggpairs' function from GGally package to create a 'pairwise' plot i.e. every variable against every other variable.
- 'ggpairs' also handles categorical data.

```{r load_ggally}

library(GGally)
```

```{r use_ggally}

ggpairs(iris, aes(colour=Species, alpha=0.7))
```
